### 2018-06-11 PAPER/main.tex
added pseudocode for AHC

\documentclass[oribibl]{llncs}
\input{includes/preamble.tex}

\begin{document}

\title{Heuristics for the Score-Constrained Strip-Packing Problem}
\author{Asyl L. Hawa \and Rhyd M. R. Lewis \and Jonathan M. Thompson}
\institute{School of Mathematics, Cardiff University, Senghennydd Road, Cardiff, UK, CF24 4AG}
\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}
\label{sec:intro}
Firstly, let us introduce the Constrained Ordering Problem (COP):

\begin{definition}
	\label{defn:cop}
	Let $\mathcal{M}$ be a multiset of unordered pairs of positive integers $\mathcal{M} = \{\{a_1, b_1\}, \{a_2,b_2\},...,\{a_n,b_n\}\}$, and let $\mathcal{T}$ be an ordering of the elements of $\mathcal{M}$ such that each element is a tuple. The Constrained Ordering Problem (COP) consists of finding a solution $\mathcal{T}$ such that, given a fixed value $\tau \in \mathbb{Z}^{+},$
	\begin{equation}
		\label{eqn:vsc}
		\emph{\textbf{rhs}}(i) + \emph{\textbf{lhs}}(i+1) \geq \tau \hspace{5mm} \forall \hspace{1mm} i \in \{1,2,..., n-1\},
	\end{equation}
	where \emph{\textbf{lhs}($i$)} and \emph{\textbf{rhs}($i$)} denote the left- and right-hand values of the $i$th tuple. The inequality is referred to as the \emph{vicinal sum constraint}.
\end{definition}

For example, given the multiset $\mathcal{M} = \{\{1,2\}, \{1,7\}, \{2,4\}, \{3,5\}, \{3,6\}, \{4,4\}\}$ and $\tau = 7$, one possible solution is $\mathcal{T} = \langle(1,2), (6,3), (5,3), (4,4), (4,2), (7,1) \rangle$.

A prominent application of the COP is in a strip-packing problem brought to light as an open-combinatorial problem by \cite{goulimis2004}. Here, a set $\mathcal{I}$ of rectangular items (where $|\mathcal{I}| = n$) of equal height $H$ made from cardboard are to be packed onto a strip of height $H$ from left to right. Each item $i \in \mathcal{I}$ has width $w_i \in \mathbb{Z}^{+}$, and possesses two vertical score lines, marked in predetermined places. A pair of knives mounted onto a bar cuts along the score lines of two adjacent items simultaneously, which allows the items to be folded with ease (see Figure \ref{fig:itemsknife}). However, by design, the scoring knives cannot be placed too close to one another, and as such have a ``minimum scoring distance'' (around 70mm in industry). The distances between each score line and the nearest edge on an item $i \in \mathcal{I}$ are the score widths, $a_i, b_i \in \mathbb{Z}^{+}$ where $a_i + b_i < w_i$, assigned such that $a_i \leq b_i$. Since these score widths are not necessarily equal, an item $i$ can be positioned in one of two orientations: ``regular'', denoted $(a_i, b_i)$, or ``rotated'', denoted $(b_i, a_i)$, where the smaller of the two score widths $a_i$ is on the left- and right- hand side of the item respectively.

\begin{figure}[H]	
	\centering
	\includestandalone[width=0.8\textwidth]{figures/itemsknife}
	\caption{Dimensions of an item $i \in \mathcal{I}$ in a regular orientation $(a_i, b_i)$, and a feasible alignment of two items that can be scored simultaneously.}	
	\label{fig:itemsknife}
\end{figure}

Clearly, for two items to be placed alongside one another feasibly, the sum of the adjacent score widths must equal or exceed the minimum scoring distance, else the knives will not be able to score the items in the correct locations. Thus, the problem consists of finding a suitable ordering and orientation of the items such that the sum of every pair of adjacent score widths is greater than or equal to the minimum scoring distance. The left-hand score width of the first item and the right-hand score width of the last item on the strip are not adjacent to any other item, and are therefore ignored.

It can be seen that this problem is in fact equal to the COP, where each unordered pair in an instance $\mathcal{M}$ contains the score widths of an item, and $\tau$ is the minimum scoring distance. It then follows that the vicinal sum constraint corresponds to the requirement for the sum of adjacent score widths to exceed $\tau$. Figure \ref{fig:items/aligned} shows the example instance $\mathcal{M}$ of the COP from above depicted as a packing problem.

\begin{figure}	
	\centering
	\begin{subfigure}[h]{0.44\textwidth}
		\includestandalone[width=\textwidth]{figures/items}
		\label{fig:items} \vspace{5pt}
	\end{subfigure} 
	\begin{subfigure}[h]{0.7\textwidth}
		\includestandalone[width=\textwidth]{figures/itemsaligned}
		\label{fig:itemsaligned}
	\end{subfigure}
	\caption{Example of the strip-packing problem, where the sum of each pair of adjacent score widths is greater than or equal to $\tau = 7$. From instance $\mathcal{M}$ above.}
	\label{fig:items/aligned}
\end{figure}

Observe that in this particular strip-packing problem the widths of the individual items are disregarded, since the aim is to arrange the items onto a strip of seemingly infinite width. However, in industrial applications, strips of material are often provided in fixed widths. Given a large problem instance, multiple strips may therefore be required to feasibly accommodate all of the items. For this reason, a more generalised problem can also be formulated.

\begin{definition}
	Let $\mathcal{I}$ be a set of rectangular items of height $H$ with varying widths $w_i \in \mathbb{Z}^+$ and score widths $a_i, b_i \in \mathbb{Z}^+$ for each item $i \in \mathcal{I}$. Given a minimum scoring distance $\tau \in \mathbb{Z}^+$, the Score-Constrained Strip-Packing Problem (SCSPP) consists of finding the minimum number of strips of height $H$ and width $W$ required to pack all $n$ items in $\mathcal{I}$ such that the sum of every pair of adjacent score widths is greater than or equal to $\tau$, and that no strip is overfilled.
\end{definition}

Note that in the special case of $\tau = 0$, the SCSPP is equivalent to the classical one-dimensional bin-packing problem (BPP). It follows that the SCSPP is at least as hard as the BPP, which is known to be NP-hard (\citealp{garey1979}), and so (under the assumption that $P \neq NP$) there is no known algorithm that is able to find an optimal solution for every instance of the SCSPP in polynomial time. Instead, heuristics can be used to find near-optimal solutions in a shorter amount of time.

\begin{definition}
	\label{defn:subprob}
	Let $\mathcal{I}' \subseteq \mathcal{I}$ be a set of items whose total width is less than or equal to the capacity of a strip, (i.e. $\sum_{i\in \mathcal{I}'}w_i \leq W$). Given a minimum scoring distance $\tau$, the Score-Constrained Packing Sub-Problem (SCPSP) involves finding an arrangement of the items in $\mathcal{I}'$ such that the sum of every pair of adjacent score widths is greater than or equal to $\tau$.
\end{definition}

The remainder of this article is structured as follows: In Section 2, we will detail an exact polynomial-time algorithm for the COP, if a solution exists. Section 3 then describes three heuristics that can be used to find feasible solutions to the SCSPP and the associated advantages and disadvantages. A comparison of these heuristics and an analysis of the results will be provided in Section 4, and finally Section 5 concludes the paper and proposes some potential directions for future work.

\section{Solving the COP}
\label{sec:ahca}
We shall now present an exact polynomial-time algorithm for the COP. The underlying algorithm was originally created by \cite{becker2010}, and is used to determine whether or not a solution exists for a given instance. In our case, we have extended this algorithm so that, if a solution does indeed exist, the algorithm is able to formulate and present us with this final solution. This is especially useful for problems such as the previously mentioned strip-packing problem, where a solution is required in order to place the items in the correct order and orientation.

Let $\mathcal{M}$ be an instance of the COP of cardinality $n$. It is now useful to model $\mathcal{M}$ a a graph $G$ in which each vertex is associated with a single value in $\mathcal{M}$ in non-decreasing order. A pair of vertices, called ``dominating vertices'' are also added to the graph, both of which are assigned values equal to $\tau$. These dominating vertices aid the solution process, and are removed at the end. Thus, the graph $G$ has $2n+2$ vertices.

As we have seen, the values in $\mathcal{M}$ are arranged in pairs. This is represented in $G$ by adding a set of ``blue'' edges, $B$, that contains edges between vertices that are ``partners'', i.e. whose values make up a pair in $\mathcal{M}$, as shown in Figure \ref{fig:partners}. By introducing a bijective function $p: V \to V$ that associates each vertex with its partner, $p(v_i) = v_j$, we can denote this set of edges as $B = \{(v_i, p(v_i)) : v_i \in V\}$. It can be seen that $B$ is a perfect matching in $G$, with $|B| = n+1$.

Next, a set of ``red'' edges $R$ is added to $G$ that consists of edges between vertices whose sum equal or exceed $\tau$, provided they are not partners. This corresponds to the vicinal sum constraint in Defintion \ref{defn:cop}. Thus, we have an undirected graph $G$ with vertex set $V = \{v_1, ..., v_{2n+2}\}$ and two distinct edge sets $B$ and $R$ such that $B \cap R = \emptyset$. Figure \ref{fig:partners/threshold} illustrates and example construction of $G$.
% Equation $v_i + v_j \geq \tau$ provided $p(v_i) \neq v_j$, i.e. not partners
% Values are assigned from $\mathcal{M}$ in non-decreasing order, i.e. value assoc with v_1 \leq value assoc with v_2 etc..
% "This corresponds to the vsc in Defn 1" - values that can be aligned alongside one another that are not part of the same unordered pair in M, hence why they cannot be partners on the graph G, must be from different unordered pairs.

\begin{figure}	
\centering
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/partners}
	\caption{$G = (V, B)$}	
	\label{fig:partners}
\end{subfigure} \hspace{40pt}
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/threshold}
	\caption{$G=(V, B \cup R)$}	
	\label{fig:threshold}
\end{subfigure}
\caption{Construction of $G$ using instance $\mathcal{M} = \{\{1,2\}, \{1,7\}, \{2,4\}, \{3,5\}, \{3,6\}, \{4,4\}\}$ and $\tau = 7$, where $v_{13}$ and $v_{14}$ are the dominating vertices.}
\label{fig:partners/threshold}
\end{figure}

Recall that a Hamiltonian cycle on a graph $G$ is a cycle that visits every vertex exactly once. Now, consider the following definition describing a variant of the classical Hamiltonian cycle involving multiple edge sets.
\begin{definition}
	\label{defn:althamcycle}
	Let $G(V, B\cup R)$ be an undirected graph, where each edge is a member of one of two sets, $B$ or $R$. $G$ contains an alternating Hamiltonian cycle if there exists a Hamiltonian cycle such that successive edges alternate between sets $B$ and $R$.
\end{definition}

An alternating Hamiltonian cycle on $G$ correpsonds to a feasible solution, as each ``blue'' edge from $B$ represents a pair of values in $\mathcal{M}$, and each ``red'' edge from $R$ shows values that can be placed alongside one another feasibly. Thus, by removing the dominating vertices and any incident edges, the remaining alternating Hamiltonian path gives the order in which the elements should be placed to produce a feasible solution $\mathcal{T}$, if indeed one exists.

Also, note that every edge in $B$ must be present in the alternating Hamiltonian cycle. Consequently, the task can be seen as finding a matching $R' \subseteq R$ of cardinality $n+1$ such that the edge sets $B$ and $R'$ form an alternating Hamiltonian cycle as described in Definition \ref{defn:althamcycle}.

To find a matching $R' \subseteq R$, the following \textsc{Maximum Cardinality Matching} (MCM) algorithm can be executed, as suggested by \cite{mahadev1994} and \cite{becker2010}. Take each vertex $v_1, v_2,..., v_{2n+2}$ in turn, and select the edge in $R$ connecting $v_i$ to the highest-indexed vertex $v_j$ that is not already incident to an edge in $R'$. Now add this edge $(v_i, v_j)$ to $R'$, and proceed to the next vertex until all the vertices have been assessed. The two vertices incident to each matching edge in $R'$ are now referred to as being ``matched''. Similarly to partners, let $m : V \to V$ be a bijective function that assigns each vertex with its match, $m(v_i) = v_j$. Then, we can denote this matching set as $R' =\{(v_i, m(v_i)) : v_i \in V\}$.

During MCM, if a vertex $v_i$ is not adjacent to any other unmatched vertex except its partner $p(v_i)$, the preceeding vertex $v_{i-1}$ can be ``rematched'', provided that 
\begin{enumerate*}[label={(\alph*)}]
	\item $v_i$ is not the first vertex; 
	%\item $v_{i-1} \in R'$; and
	\item $(v_{i-1}, m(v_{i-1})) \in R'$; and
	\item $(v_{i-1}, p(v_i)) \in R$.
\end{enumerate*}
Then, $v_i$ is matched with the vertex that is currently matched with $v_{i-1}$, i.e $m(v_i) = m(v_{i-1})$, and $v_{i-1}$ is rematched with $p(v_i)$, i.e. $m(v_{i-1}) = p(v_i)$. It is guaranteed that MCM will produce a maximum cardinality matching $R' \subseteq R$, and is able to do so in at most $O(n \log n)$ time (\citealp{mahadev1994}).

\begin{figure}	
	\centering
	\begin{subfigure}[h]{0.4\textwidth}
		\includestandalone[width=\textwidth]{figures/matching}
		\caption{$G' = (V, B \cup R')$}	
		\label{fig:matching}
	\end{subfigure} \quad
	\begin{subfigure}[h]{0.55\textwidth}
		\includestandalone[width=\textwidth]{figures/mps}
		\caption{Planar representation of $G' =(V, B \cup R')$}	
		\label{fig:mps}
	\end{subfigure}
	\caption{Spanning subgraph $G'$ consisting of a matching set $R' \subseteq R$ derived using MCM. The same solution is shown in planar form in (b).}
	\label{fig:matching/mps}
\end{figure}

Clearly, if $|R'| < n+1$ after MCM has completed, there are insufficient red edges to form an alternating Hamiltonian cycle, and hence no feasible solution exists for the given instance $\mathcal{M}$ of the COP. Otherwise, $R'$ is a perfect matching of cardinality $n+1$ and the spanning subgraph $G'=(V, B \cup R')$ is a 2-regular graph, where each vertex $v_i \in V$ is adjacent to its partner $p(v_i)$ via a blue edge in $B$, and its match $m(v_i)$ via a red edge in $R'$ (see Figure \ref{fig:matching/mps}). $G'$ therefore consists of a collection of cyclic components $C_1, C_2, ..., C_q$. In the case where $G'$ comprises one component, i.e. $q = 1$, then this alternating cycle is in fact Hamiltonian, and therefore specifies a solution to the COP. However, if $q > 1$, the components of $G'$ must be combined to form a single alternating Hamiltonian cycle. To do this, a \textsc{Bridge Recognition} (BR) algorithm can be executed that selects suitable edges from $R$ that will replace certain edges in $R'$ in order to function as bridges and connect the components of $G'$.

BR operates by first ordering the edges in $R'$ such that the lower-indexed vertices of each edge are in increasing order and the higher-indexed vertices are in decreasing order, i.e. $(v_1, v_{2n+2}), (v_2, v_{2n+1}), ..., (v_{n+1}, v_{n+2})$. Any edges that cannot be placed in such an order are then removed from this list. For instance, in the example illustrated in Figure \ref{fig:matching/mps}, the edges would be sorted as follows: $(v_1, v_{14}), (v_2, v_{13}),(v_3, v_{12}),(v_4, v_{11}),(v_6, v_{10}),(v_7, v_8)$.
% (v5, v9) removed from list, not matched with highest-indexed vertex in MCM since it was v10 and v10 is v5's partner.

Starting from the first edge in the list, BR then searches through the list to find an edge that meets the following conditions: 
\begin{enumerate*}[label={(\alph*)}]
	\item The lower-indexed vertex of the current edge and the higher-indexed vertex of the next edge in the list are adjacent via an edge in $R$; and
	\item The current edge and the next edge are members of different components on $G'$.
\end{enumerate*}
If these conditions are met, BR adds the current edge to a set $R_i$, and continues to add all succeeding edges in the list to $R_i$ provided that, for each edge, both conditions hold and the succeeding edge is not a member of a component of $G'$ that has an edge in $R_i$. Once there are no more valid edges available to add to $R_i$, BR resumes its search through the remaining edges in the list to find a new edge that meets the conditions, and can begin a new set $R_{i+1}$. The procedure terminates once the penultimate edge in the list has been assessed.
% (a) (v_i,m(v_{i+1})) \in R
% (b) (v_i, m(v_i)) \in C_k, (v_{i+1}, m(v_{i+1})) \in C_l, k \neq l

Now, one of the following three conditions holds.
\begin{enumerate}
\item In the event that BR has been unable to produce a single set, there are no suitable edges that can combine the components $C_1, C_2, ..., C_q$ of $G'$, and therefore an alternating Hamiltonian cycle for $G$ cannot be created. Thus, no feasible solution exists for the corresponding COP.

\item On the other hand, if there exists a set $R_i$ such that $|R_i| = q$, then the components of $G'$ can be merged together to form a single alternating Hamiltonian cycle. This is achieved by adding the red edge connecting the lower-indexed vertex of each edge in $R_i$ to the higher-indexed vertex of the next edge to $R'$ (for the final edge in $R_i$, add the red edge connecting its lower-indexed vertex to the higher-indexed vertex of the first edge). Any edges that appear in both $R_i$ and $R'$ are then removed from $R'$, so that $R'$ remains a perfect matching and $R_i \cap R' = \emptyset$. $G'$ then consists of a single alternating Hamiltonian cycle, and hence a solution has been found.
% add edges (v_i, m(v_{i+1})) for i \in {1, 2, ..., |R_i| -1} from R to R', and then the edge (v_{|R_i|}, m(v_1)) from R to R'

\item Otherwise, it may be that multiple sets are required to connect the components of $G'$. For two sets $R_i$ and $R_j$ to ``overlap'', each set must have one edge from the same component in $G'$, with the other edges in each set being from different components. For example, if $R_1$ has two edges, one from component $C_1$ and one from component $C_2$, and $R_2$ has three edges that are members of components $C_2$, $C_3$ and $C_4$, then $R_1$ and $R_2$ are said to overlap, since they each possess an edge $C_2$, but the other edges in the sets are from different components. Hence, a selection of sets need to be found such that each set overlaps with at least one other set, and each component has at least one edge in one of the selected sets. If such a collection of sets exists, then the components of $G$ can be merged to create an alternating Hamiltonian cycle by applying the connecting procedure above to every set in the collection.
% Refer to figures
\end{enumerate}
Using the instance illustrated in the figures as an example, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are in the set $R_1$ formed by BR, as they are members of different components ($(v_3, v_{12}) \in C_1$ and $(v_4, v_{11}) \in C_2$ as seen in Figure \ref{fig:mps}), and $(v_3, v_{11}) \in R$. Note that since $v_4$ and $v_{11}$ are adjacent, $v_4$ must also be adjacent to $v_{12}$, i.e. $(v_4, v_{12}) \in R$, since the value associated with $v_{12}$ is greater than or equal to the value associated with $v_{11}$. Then, as $|R_1| = q$, the edges $(v_3, v_{12})$ and $(v_4, v_{11})$ are removed from $R'$ and replaced by the edges $(v_3, v_{11})$ and $(v_4, v_{12})$ from $R$, as shown in Figure \ref{fig:connect/cycle}. Removing the dominating vertices and any incident edges results in an alternating Hamiltonian path, which corresponds to a feasible solution $\mathcal{T}$.

This algorithm, which will be referred to as the \textsc{Alternating Hamiltonian Construction} (AHC) algorithm is able to deduce whether a feasible solution $\mathcal{T}$ exists for any given instance $\mathcal{M}$ of the COP. If a solution does indeed exist, AHC can produce this solution in at most $O(n^2)$ time.
% Cite Becker

\begin{figure}	
	\centering
	\begin{subfigure}[H]{0.7\textwidth}
		\includestandalone[width=\textwidth]{figures/mpsconnect}
		\caption{Edges from $R\backslash R'$ can connect vertices from different cycles.}	
		\label{fig:mpsconnect} \vspace{5mm}
	\end{subfigure}
	\begin{subfigure}[H]{0.7\textwidth}
		\includestandalone[width=\textwidth]{figures/mpscycle}
		\caption{Replacing edges connects the components and forms an alternating Hamiltonian cycle.}	
		\label{fig:mpscycle}
	\end{subfigure}
	\caption{Selected edges found using BR merge the components of $G'$ together, forming an alternating Hamiltonian cycle which corresponds to a feasible solution $\mathcal{T} = \langle(1,2), (6,3), (5,3), (4,4), (4,2), (7,1) \rangle$.}
	\label{fig:connect/cycle}
\end{figure}

\begin{algorithm}[h]
\caption{\textsc{Alternating Hamiltonian Construction} (AHC) algorithm}
\label{alg:ahca}
\begin{algorithmic}[1]
	\State Create the graph $G = (V, B \cup R)$ from instance $\mathcal{M}$
	\State Execute \textsc{Maximum Cardinality Matching}
	\If{$|R'| < n + 1$}
		\State Infeasible, end
	\EndIf
	\State Produce the subgraph $G' = (V, B \cup R')$
	\If{$q = 1$}
		\State Remove dominating vertices and incident edges
		\Return $G'$
	\Else
		\State Execute \textsc{Bridge Recognition}
		\If{$q = 1$}
			\State Remove dominating vertices and incident edges
			\Return $G'$
		\Else
			\State Infeasible, end	
		\EndIf	
	\EndIf
\end{algorithmic}	
\end{algorithm}

\begin{theorem}
	\label{thm:copsoln}
	If there exists an alternating Hamiltonian cycle on a graph $G$ created from an instance $\mathcal{M}$ of the COP, then AHC is able to find the cycle in $O(n^2)$ time.
\end{theorem}
\begin{proof}
	\qed
\end{proof}

\section{Heuristics for the SCSPP}
\label{sec:scsppsoln}

In this section, we now turn our attention to the multi-strip version of the problem. As mentioned in the introduction, this problem is a generalisation of the BPP in that we also require the sum of every pair of adjacent score widths to be greater than or equal to a minimum scoring distance $\tau$. It follows that the SCSPP is at least as hard as the BPP, which is known to be NP-hard (\citealp{garey1979}), and so (under the assumption that $P \neq NP$) there is no known algorithm that is able to find an optimal solution for every instance of the SCSPP in polynomial time. Instead, heuristics can be used to find near-optimal solutions in a shorter amount of time.

For an instance of the SCSPP, a feasible solution is represented by the set $\mathcal{S} = \{S_1, S_2, ..., S_k\}$ such that
\begin{subequations}
	\begin{align}
		\bigcup\nolimits_{j=1}^{|\mathcal{S}|} S_j &= \mathcal{I}, \label{eqn:packall}\\[3pt]
		S_j \cap S_l &= \emptyset \hspace{5mm} \forall \hspace{1mm} j, l \in \{1, 2, ..., |\mathcal{S}|\}, \hspace{2mm} j \neq l, \label{eqn:nooverlap} \\[3pt]
		\sum\nolimits_{i=1}^{|S_j|}w_i & \leq W \hspace{3mm} \forall \hspace{1mm} S_j \in \mathcal{S}, \label{eqn:capacity} \\[3pt]
		\textup{\textbf{rhs}}(i) + \textup{\textbf{lhs}}(i+1) &\geq \tau \hspace {5mm} \forall \hspace{1mm} i \in \{1, 2, ..., |S_j|-1\}, \hspace{2mm} \forall \hspace{1mm} S_j \in \mathcal{S}. \label{eqn:vscstrip}
\end{align}
\end{subequations}

That is, all items must be packed onto a strip (\ref{eqn:packall}), each item can only be placed on one strip (\ref{eqn:nooverlap}), the strips cannot be overfilled (\ref{eqn:capacity}), and the items on each strip $S_j$ in the solution must be arranged such that the vicinal sum constraint is fulfilled, that is, the sum of every pair of adjacent score widths is greater than or equal to the minimum scoring distance $\tau$ (\ref{eqn:vscstrip}).

An optimal solution $\mathcal{S}$ for the SCSPP is a solution that consists of the fewest number of strips $k$ needed to feasibly contain the $n$ items in the given problem instance. A simple lower bound for $k$ is the theoretical minimum
\begin{equation}
\label{eqn:lowerbound}
t = \ceil*{\frac{\sum_{i=1}^{n}w_i}{W}},
\end{equation}
which can be computed in $O(n)$ time (\citealp{martello1990b}). Perhaps the simplest and most well-known heuristic for packing problems is \textsc{First-Fit} (FF), a greedy online algorithm that places each item presented in some arbitrary order onto the lowest-indexed strip such that the capacity of the strip is not exceeded. It is known that there always exists at least one ordering of the items such that FF produces an optimal solution (\citealp{lewis2009}). A minor modification to FF yields the \textsc{First-Fit Decreasing} (FFD) heuristic, which sorts the items in non-increasing order of size prior to performing FF. In 1973, \citeauthor{johnson1973} showed that FFD is guaranteed to return a solution that uses no more than $\frac{11}{9}k + 4$ strips. More recently, \cite{dosa2007} has proven that the worst case for FFD is in fact $\frac{11}{9}k + \frac{6}{9}$, and that this bound is tight. Due to the initial sorting of the items in non-increasing order of sizes, the time complexity of FFD is $O(n\log n)$.

As mentioned in the introduction, the SCSPP shares many similarities with the BPP, however the addition of constraint (\ref{eqn:vscstrip}) induces complications that must be taken into consideration. One obvious difference is the order in which the items must appear on the strips. The order of the items in the BPP is unimportant, as there are no conditions regarding which items can be placed together, however in the SCSPP the items must be ordered in a way that meets the vicinal sum constraint. In addition, removing an item in the BPP retains feasibility, whereas in the SCSPP this is not guaranteed, as it may leave a subset of items for which the vicinal sum constraint is not satisfied. Furthermore, the theoretical minimum $t$ (Equation (\ref{eqn:lowerbound})) has the potential to be less accurate for the SCSPP, as the minimum scoring distance $\tau$ is not considered. For example, if the minimum scoring distance is greater than twice the largest score width, then it is clear that $n$ strips will be required, regardless of the items' widths.
% Time complexity for FFD also inaccurate

Three heuristics for the SCSPP have been developed: a basic FFD heuristic with a simple modification; a heuristic that packs strips individually and prioritises score widths; and a powerful version of FFD that incorporates the polynomial-time AHC algorithm.

\subsection{Modified First-Fit Decreasing Heuristic (MFFD)}
\label{sec:mffd}
The \textsc{Modified First-Fit Decreasing} (MFFD) heuristic performs in the same manner as the original FFD, attempting to place each item onto the end of the lowest-indexed strip. If an item is able to be packed onto a strip without exceeding the strip's capacity, MFFD then checks to see if the vicinal sum constraint is met between the right-most score on the strip and one of the score widths on the current item. If the constraint is met, MFFD places the item on the end of the strip in the appropriate orientation, else the heuristic attempts to pack the item onto the next lowest-indexed strip. For example, in Figure \ref{fig:comparestrips}, the strip is able to accommodate the current item, however the vicinal sum constraint is not fulfilled in either orientation of the item, and therefore it cannot be feasibly packed onto the strip.

Despite the additional step, MFFD is a very fast heuristic, as checking the vicinal sum constraint can be performed in constant time. The most prominent issue with this heuristic, however, is due to the items being placed on the end of the strips. Although an item may not meet the constraint when placed on the end of a strip, there may be another location on the strip where the item can be placed feasibly. MFFD is incapable of placing items anywhere other than on the end of the strips, and so an item that could potentially be packed must be placed on another strip, or perhaps even begin a new strip, thus increasing the number of strips in the final solution.

\subsection{Pair-Smallest Heuristic (PS)}
\label{sec:ps}

The \textsc{Pair-Smallest} (PS) heuristic is an extension of a procedure defined by \cite{lewis2011}. Unlike MFFD, which packs each item in turn, PS focusses on packing one strip at a time, only starting a new strip once the current strip is unable to accommodate any more items feasibly. Each strip is initialised by choosing the item from $\mathcal{I}$ with the smallest score width, and packing it in a regular orientation. PS then continues to fill the strip by selecting the item with the smallest score width that meets the vicinal sum constraint with the right-most score width on the strip, and whose width will not cause the strip to be overfilled. This heuristic aligns the smallest score widths with the largest ones, eliminating the possibility of placing larger score widths together unecessarily. Note that PS prioritises the vicinal sum constraint over the item widths, choosing to fulfil the constraint first before considering whether the item can be accommodated. If there are multiple items with equal score widths, break ties by choosing the item whose second score width is the largest.
% Break ties by choosing item whose second score width is larger.
% If item with smallest feasible score width does not fit onto the strip, choose item with next smallest score width and repeat.

\subsection{Modified First-Fit Decreasing with AHC (MFFD$^+$)}
\label{sec:mffd+}

The last heuristic incorporates AHC as described in Section \ref{sec:ahca}. The \textsc{Modified First-Fit Decreasing with AHC} (MFFD$^+$) heuristic operates in a similar fashion to the MFFD, placing items sorted in decreasing order onto the lowest-indexed strip. However, when the lowest-indexed strip is found that can accommodate the current item, MFFD$^+$ then executes AHC on the sub-problem, where the items on the strip and the current item form the subset $\mathcal{I}' \subseteq \mathcal{I}$ (see Definition \ref{defn:subprob}). If AHC finds a feasible solution, the items are placed on the strip in the order of the solution, which includes the current item, else MFFD$^+$ attempts to pack the current item on the next lowest-indexed strip. Using AHC means that if a feasible alignment of the items exists, there is a guarantee that it will be found. Unlike MFFD, where the current item can only be placed on the end of the strip, MFFD$^+$ allows the items to be entirely rearranged (see Figure \ref{fig:comparestrips}). This reduces the possibility of having to start a new strip for an item, thus preventing increasing the number of strips in the final solution. Since AHC is executed every time an item can fit onto a strip that already contains items, (up to $|\mathcal{S}|$ times per item, where $|\mathcal{S}|$ is the number of strips containing at least one item so far in the procedure) MFFD$^+$ is slower in comparison to the previous heuristics.

\begin{figure}[h!]	
	\centering
	\includestandalone[width=\textwidth]{figures/strips}
	\caption{Example instance of the sub-problem, with $W = 20$ and $\tau = 70$. In MFFD the constraint is not fulfilled in either orientation, however MFFD$^+$ is able find a feasible arrangement.}	
	\label{fig:comparestrips}
\end{figure}

\section{Comparison}
For our experiments, we generated 1000 problem instances for two different sizes of $\mathcal{I}$ containing 500 and 1000 items. For each problem instance the items have varying widths $w_i$ between 150 and 1000 and score widths $a_i, b_i$ between 1 and 70 to ensure that each item has exactly two score lines, all selected randomly from a uniform distribution. Strips of widths $W = 5000$, 2500 and 1250 were used to alter the number of items per strip. Both the items and the strips have equal height of $H = 1$. Similarly to experiments performed by \cite{lewis2011}, we also introduced a parameter $\delta$ to denote the proportion of pairs of score widths from different items that meet the vicinal sum constraint, i.e. whose sum is greater than or equal to $\tau$. Values of $\delta$ from 0.0 to 1.0 were created by changing the value of the minimum scoring distance $\tau$. Clearly, when $\delta = 0.0$, there are no score widths that can be packed together feasibly, and so $n$ strips will be required (one for each item), whereas if $\delta = 1.0$ all pairs of score widths meet the constraint, and the problem is analogous to the BPP. (As the width of the strips decrease, the average number of items per strip also decreases, making the problem more difficult.)

The heuristics were implemented in C++ and executed on a computer equipped with an Intel Core i3-2120 processor with a frequency of 3.30GHz. The source code for these procedures used for our computational experiments is available at \cite{hawa2018}. We calculated the average optimality of each heuristic for every combination of $n$, $\delta$, and $W$ from 1000 instances, where the optimality $r$ of each instance is
\begin{equation}
	r = \frac{|\mathcal{S}|}{t},
\end{equation}
i.e. the ratio of the number of strips in the final solution $\mathcal{S}$ generated by the heuristic to the theoretical minimum $t$, which was calculated using Equation \ref{eqn:lowerbound}. All individual trials were executed in under 160ms.

Tables \ref{table:n=500} and \ref{table:n=1000} compare the results obtained from the three heuristics using different values of $\delta$ and $W$ for $n = 500$ and 1000 respectively. Figures in bold indicate the best average optimality for the given combination of parameters. Initially, we can see that $r$ tends toward 1 as $\delta$ increases, since the proportion of score widths meet the vicinal sum constraint increases and more items can be packed onto one strip, reducing the number of strips required. Note that when $\delta = 1.0$ MFFD and MFFD$^+$ have identical optimality ratios, as the instances are equivalent to the original BPP ($\tau = 0$) and so they operate in the same fashion as the original FFD heuristic.

Looking at results for both $n = 500$ and $n = 1000$, there is a clear pattern with respect to $r$ across all widths and proportion levels. PS has the best optimality ratio for a wider range of $\delta$ when the strips are wider, and a much smaller range when the strip width decreases. Conversely, MFFD$^+$ obtains better optimality ratios for a wider range of $\delta$ when $W = 1250$. Although PS does have the best optimality of the three heuristics for $\delta = 0.1$ and 0.2 using the smallest strip width, we can see that it only marginally superior to MFFD$^+$. For example, take $\delta = 0.2$ in Table \ref{table:n=500} for $W = 1250$. The difference between $r$ for PS and MFFD$^+$ is 0.00829, which translates to less than 2 strips difference between the average number of strips generated by each heuristic.

One interesting feature is the reduction of $r$ between $\delta = 0.4$ and $\delta = 0.5$, which is much more significant for $W = 5000$ as opposed to $W = 1250$. The reduction occurs due to an increase in the number of feasible alignments of the items, as every score width is able to be paired with $n-1$ other score widths when $\delta = 0.5$. Thus, when $W = 5000$, more items can be packed onto a strip feasibly, resulting in fewer strips in the final solution. However, the difference in $r$ is less noticeable for $W = 1250$, as regardless of the increased proportion in possible pairings, the capacity of the strips is much smaller, preventing the collection of items from multiple strips onto one.

Although using the average optmality from 1000 instances provides a useful overview of the efficiency of a heuristic, there are other characteristics that we can consider. Take, for example, the results obtained with parameters $\delta = 0.7$ and $W = 5000$ in Table \ref{table:n=1000}. Clearly PS obtains solutions with the fewest strips on average, however, we noted that out of the 1000 instances, PS did not produce a single solution containing $t$ strips. On the other hand, there were 152 instances in which MFFD$^+$ was able to generate a solution $S$ such that $|\mathcal{S}| = t$, thus implying that there are at least 152 solutions that can be solved to optimality. Despite this, MFFD$^+$ has a higher average optimality than PS, suggesting that the variance in the number of strips required is higher.

\begin{enumerate}
	\item MFFD and MFFD$^+$ able to obtain opts $\leq 1\%$ of $t$, PS is not.
	\item PS never best $\delta \geq 0.8$.
	\item Significant difference in opt, $n = 1000$, $W = 1250$, $\delta = 0.6$, difference between MFFD$^+$ and PS opts is 0.10912 (MFFD$^+$ better), which amounts to an average difference of over 50 strips in solns produced.
	\item anomaly MFFD $n = 1000$, $W = 1250$, $\delta = 0.3$
	\item harder problems, \cite{falkenauer1996}, triplets, fewer items per strip, further from 1 when shorter strips.
	\item cases when opt $\neq 1$: $t$ incorrect, solution produced is not optimal, or both.
	\item less than 10\% past $\delta = 0.8$
\end{enumerate}

\begin{table}[!htb]
	\centering
	\ra{1.3}
	\caption{Average optimality ratios for $n = 500$. The average number of items per strip for $W = 5000$, 2500 and 1250 are 8.4746, 4.3103, and 2.1645 respectively.}
	\begin{tabular}{@{}c@{\hspace{25pt}}c@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}c@{}}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 58.039$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 2500$, $t = 115.571$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 1250$, $t = 230.648$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.61813 & 8.61813 & 8.61813 && 4.32789 & 4.32789 & 4.32789 && 2.16858 & 2.16858 & 2.16858 \\
		0.1 & 5.21367 & \textbf{4.84177} & 5.16085 && 2.65934 & \textbf{2.47735} & 2.65679 && 1.51486 & \textbf{1.47896} & 1.51486 \\
		0.2 & 4.03055 & \textbf{3.45946} & 3.97645 && 2.12099 & \textbf{1.84737} & 2.11788 && 1.32613 & \textbf{1.31781} & 1.32610 \\
		0.3 & 3.11091 & \textbf{2.34803} & 3.03816 && 1.72955 & \textbf{1.39676} & 1.70772 && 1.19451 & 1.22922 & \textbf{1.19381} \\
		0.4 & 2.43605 & \textbf{1.52907} & 2.29713 && 1.45976 & \textbf{1.12798} & 1.40951 && 1.11037 & 1.18065 & \textbf{1.10819} \\
		0.5 & 1.91057 & \textbf{1.04130} & 1.69129 && 1.26343 & \textbf{1.03341} & 1.19575 && 1.05771 & 1.15415 & \textbf{1.05319} \\
		0.6 & 1.49086 & \textbf{1.01263} & 1.24645 && 1.12448 & \textbf{1.02957} & 1.06874 && 1.02932 & 1.13470 & \textbf{1.02398} \\
		0.7 & 1.19571 & \textbf{1.01257} & 1.04468 && 1.04933 & 1.02774 & \textbf{1.01918} && 1.01741 & 1.11368 & \textbf{1.01444} \\
		0.8 & 1.04977 & 1.01203 & \textbf{1.00815} && 1.01595 & 1.02661 & \textbf{1.00790} && 1.01297 & 1.09136 & \textbf{1.01197} \\
		0.9 & 1.00929 & 1.01197 & \textbf{1.00463} && 1.00734 & 1.02586 & \textbf{1.00641} && 1.01159 & 1.07283 & \textbf{1.01139} \\
		1.0 & \textbf{1.00423} & 1.01171 & \textbf{1.00423} && \textbf{1.00626} & 1.02561 & \textbf{1.00626} && \textbf{1.01130} & 1.06549 & \textbf{1.01130} \\
		\bottomrule
	\end{tabular}	
	\label{table:n=500}
\end{table}

\begin{table}[!htb]
	\centering
	\ra{1.3}
	\caption{Average optimality ratios for $n = 1000$. The average number of items per strip for $W = 5000$, 2500 and 1250 are 8.6207, 4.329 and 2.1692 respectively.}
	\begin{tabular}{@{}c@{\hspace{25pt}}c@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}c@{}}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 115.534$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 2500$, $t = 230.563$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 1250$, $t = 460.623$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.65711 & 8.65711 & 8.65711 && 4.33803 & 4.33803 & 4.33803 && 2.17138 & 2.17138 & 2.17138 \\
		0.1 & 5.17263 & \textbf{4.84248} & 5.14027 && 2.63606 & \textbf{2.48111} & 2.64271 && 1.51056 & \textbf{1.46744} & 1.51121 \\
		0.2 & 3.97565 & \textbf{3.46163} & 3.95164 && 2.09301 & \textbf{1.85824} & 2.10161 && 1.31785 & \textbf{1.31090} & 1.31876 \\
		0.3 & 3.04866 & \textbf{2.35023} & 3.01260 && 1.69777 & \textbf{1.40862} & 1.68776 && \textbf{1.18304} & 1.22111 & 1.18344 \\
		0.4 & 2.37436 & \textbf{1.52046} & 2.26560 && 1.42646 & \textbf{1.13101} & 1.38430 && 1.09722 & 1.17063 & \textbf{1.09614} \\
		0.5 & 1.84748 & \textbf{1.02600} & 1.64169 && 1.22960 & \textbf{1.03017} & 1.16950 && 1.04411 & 1.14430 & \textbf{1.03986} \\
		0.6 & 1.43331 & \textbf{1.01212} & 1.20264 && 1.09935 & \textbf{1.02738} & 1.05010 && 1.01971 & 1.12392 & \textbf{1.01480} \\
		0.7 & 1.15500 & \textbf{1.01198} & 1.02574 && 1.03421 & 1.02569 & \textbf{1.01224} && 1.01150 & 1.10164 & \textbf{1.00926} \\
		0.8 & 1.03182 & 1.01182 & \textbf{1.00587} && 1.01111 & 1.02454 & \textbf{1.00682} && 1.00861 & 1.07791 & \textbf{1.00804} \\
		0.9 & 1.00707 & 1.01165 & \textbf{1.00419} && 1.00663 & 1.02406 & \textbf{1.00608} && 1.00784 & 1.05885 & \textbf{1.00774} \\
		1.0 & \textbf{1.00399} & 1.01169 & \textbf{1.00399} && \textbf{1.00603} & 1.02400 & \textbf{1.00603} && \textbf{1.00770} & 1.05406 & \textbf{1.00770} \\
		\bottomrule
	\end{tabular}
	\label{table:n=1000}	
\end{table}

\begin{comment}
\begin{itemize}
	\item ``BPP becomes easier to solve when the number of items per bin increases, and problems with two or less items per bin are trivial to solve. Triplets are proposed to be `the most difficult bin packing instances''' \cite{falkenauer1996}.
	\item Fair comparison, running over 1000 instances (average of 1000 instances)
	\item Note that MFFD and MFFD$^+$ perform identically when $\delta = 1.0$, as they both perform in exactly the same manner as the original FFD heuristic  
	\item All instances completed in less than 160ms
	\item PS is not able to reach highest optimality in any case from $\delta = 0.8$ onwards.
	\item If optimality = 1 then lower bound reached
	\item state that bold is best optimality
	\item PS never obtains best avg optimality at  $\delta = 1.0$ in any case
	\item As W decreases, i.e. items/strip decreases, PS becomes less efficient, mffd+ better
	\item n = 1000 W = 1250 PS doesnt obtain soln $<$ 10\% for any value of $\delta$
	\item mffd+ $<$ 10\% for $\delta = 0.8$ onwards in all cases
	\item mffd+ and mffd+ able to get $<$ 10\% for W = 1250 for n = 100 and 500 $\delta = 0.5$, and n = 1000 $\delta = 0.4$ onwards
	\item PS never obtains avg solns $<$ 1\% in any case
	\item In a few cases, mffd+ able to get avg opt $<$ 1\% for smaller values of $\delta$ than mffd.
	\item PS able to get $<$ 10\% for $\delta = 0.6$ onwards in n= 100, 500, 1000, and W=5000,2500, but reduces significantly when W = 1250.
	\item First row is avg number of items per strip.
\end{itemize}
\end{comment}

\begin{comment}
\begin{table}[!htb]
	\centering
	\ra{1.2}
	\caption{Average optimality ratios for $|\mathcal{I}| = 100$. Average $t$ for $W = 5000$, 2500 and 1250 are 11.993, 23.478, and 46.466 respectively.}
	\begin{tabular}{@{}c@{\hspace{25pt}}c@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}c@{}}\toprule
		& \multicolumn{3}{c}{$W = 5000$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 2500$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 1250$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0	& 8.35709 & 8.35709 & 8.35709 && 4.26800 & 4.26800 & 4.26800 && 2.15624 & 2.15624 & 2.15624 \\
		0.1	& 5.27284 & \textbf{4.89963} & 5.14404 && 2.72299 & \textbf{2.48800} & 2.69191 && 1.53147 & \textbf{1.51049} & 1.52968 \\
		0.2	& 4.16155 & \textbf{3.46467} & 3.98645 && 2.21096 & \textbf{1.87449} & 2.16338 && 1.34812 & \textbf{1.34499} & 1.34593 \\
		0.3	& 3.27805 & \textbf{2.40153} & 3.05546 && 1.82199 & \textbf{1.43724} & 1.75610 && 1.22453 & 1.24810 & \textbf{1.21994} \\
		0.4	& 2.61641 & \textbf{1.63390} & 2.33662 && 1.54989 & \textbf{1.16887} & 1.46534 && 1.14507 & 1.19631 & \textbf{1.13979} \\
		0.5	& 2.08260 & \textbf{1.13827} & 1.77494 && 1.35101 & \textbf{1.04978} & 1.25567 && 1.09274 & 1.17069 & \textbf{1.08661} \\
		0.6 & 1.65897 & \textbf{1.02027} & 1.36093 && 1.20040 & \textbf{1.03720} & 1.11964 && 1.06255 & 1.15766 & \textbf{1.05701} \\
		0.7	& 1.32002 & \textbf{1.01350} & 1.11348 && 1.09659 & \textbf{1.03556} & 1.04569 && 1.04546 & 1.14416 & \textbf{1.04150} \\
		0.8	& 1.10725 & \textbf{1.01384} & 1.02188 && 1.03577 & 1.03439 & \textbf{1.01696} && 1.03711 & 1.12951 & \textbf{1.03565} \\
		0.9	& 1.02134 & 1.01315 & \textbf{1.00646} && 1.01364 & 1.03285 & \textbf{1.01029} && 1.03434 & 1.11631 & \textbf{1.03416} \\
		1.0 & \textbf{1.00480} & 1.01302 & \textbf{1.00480} && \textbf{1.00919} & 1.03261 & \textbf{1.00919} && \textbf{1.03393} & 1.10807 & \textbf{1.03393} \\
		\bottomrule
	\end{tabular}	
	\label{table:n=100}
\end{table}
\end{comment}

\begin{comment}
\textbf{$n$ = 100}
\begin{itemize}
	\item Average total item width = 57466.776, $\sigma$ = 2499.06092
	\item $W = 5000$, Average LB = 11.993, $\sigma$ = 0.56829, i/s = 8.3333
	\item $W = 2500$, Average LB = 23.478, $\sigma$ = 1.05618, i/s = 4.1667
	\item $W = 1250$, Average LB = 46.466, $\sigma$ = 2.02703, i/s = 2.1277
\end{itemize}

\textbf{$n$ = 500}
\begin{itemize}
	\item Average total item width = 287683.855, $\sigma$ = 5434.63046
	\item $W = 5000$, Average LB = 58.039, $\sigma$ = 1.12227, i/s = 8.4746
	\item $W = 2500$, Average LB = 115.571, $\sigma$ = 2.18150, i/s = 4.3103
	\item $W = 1250$, Average LB = 230.648, $\sigma$ = 4.35041, i/s = 2.1645
\end{itemize}

\textbf{$n$ = 1000}
\begin{itemize}
	\item Average total item width = 575163.174, $\sigma$ = 7880.93401
	\item $W = 5000$, Average LB = 115.534, $\sigma$ = 1.59463, i/s = 8.6207
	\item $W = 2500$, Average LB = 230.563, $\sigma$ = 3.16734, i/s = 4.3290
	\item $W = 1250$, Average LB = 460.623, $\sigma$ = 6.30895, i/s = 2.1692
\end{itemize}

\end{comment}
\begin{comment}
\section{Further Work}
\begin{itemize}
	\item mffd+ items are packed one by one, could be a case where an items fits onto a strip but ahc cannot find a feasible solution, so item must be packed elsewhere, but if two new items could fit on the strip, ahc might be able to find a solution. 
	\item include item width in ahc, so add values to the blue edges of the graph, weighted edges, red edges have value = 0.
	\item Then the problem would be: find the minimum number of alternating cycles on $G$ such that the length of the cycle is less than or equal to some value $W$ (which is the strip width). The length of the cycle would be the sum of all the weighted edges, so the sum of all the blue edges, since red = 0. 
	\item then could break and form edges to create new components of different sizes if some are too large, or join together if some are small, thus decreasing the number of cycles
	\item or if possible, find a single alt Ham cycle, and break up into as few smaller cycles as possible such that the length of each cycle is less than or equal to $W$
	\item another way, create as few feasible subsets $\mathcal{I}' \subseteq \mathcal{I}$ such that the total width of the items in each subset $\mathcal{I}$ is less than $W$, i.e. perform usual FFD alg, ignore score widths
	\item then perform ahc on each set, try to find feasible solution.
	\item if sets dont have feasible solution, try swapping items between sets
	\item initial tests, Becker theorem 6 2015 paper 
\end{itemize}
\end{comment}

\begin{comment}
	\begin{enumerate}
	\item Opt ratios tend to 1 as $\delta \to 1$, as the problem instances become easier to solve, more feasible alignments of pairs of items, more items per strip, reducing the need for more strips, closer to BPP.
	\item $\delta = 0.0$ $n$ strips required, hence all ratios equal.
	\item $\delta = 1.0$,  MFFD and MFFD$^+$ obtain identical opts, since (described above) instance is equivalent to original BPP, $\tau = 0$, so MFFD and MFFD$^+$ perform in the same fashion as original FFD.
	\item Pattern, PS best for wider range of $\delta$ when $W$ high, and smaller range (only 0.1-0.2) when $W = 1250$.
	\item Conversely, MFFD$^+$ best when $W$ small for wider range of $\delta$.
	\item Note that although PS best for 0.1-0.2, difference is negligible, e.g. $n = 500$, $\delta = 0.2$, difference in opt is 0.00829, translates to less than 2 strips difference on average.
	\item MFFD and MFFD$^+$ able to obtain opts $\leq 1\%$ of $t$, PS is not.
	\item PS never best $\delta \geq 0.8$.
	\item $n = 1000$, $W = 5000$, $\delta = 0.7$, PS is best opt, however PS does not produce a single solution containing $t$ strips, whereas there are 152 instances in which MFFD$^+$ produces solutions containing $t$ strips, i.e. there are at least 152 instances out of the 1000 where $k = t$, and PS cannot obtain any. Therefore, MFFD$^+$ solutions must vary significantly (i.e. some solns that are much worse than optimal), as PS is best opt.
	\item Significant difference in opt, $n = 1000$, $W = 1250$, $\delta = 0.6$, difference between MFFD$^+$ and PS opts is 0.10912 (MFFD$^+$ better), which amounts to an average difference of over 50 strips in solns produced.
	\item Noticable drop in opts between $\delta = 0.4$ and $\delta = 0.5$ when $W$ high, as capacity not main concern, more items can now be placed next to one another, so number of strips reduces dramatically, however when $W = 1250$ there is still a capacity issue, regardless of the increased proportion of possible pairings.
	\item anomaly MFFD $n = 1000$, $W = 1250$, $\delta = 0.3$
	\item harder problems, \cite{falkenauer1996}, triplets, fewer items per strip.
	\item cases when opt $\neq 1$: $t$ incorrect, solution produced is not optimal, or both.
	\item less than 10\% past $\delta = 0.8$
	\end{enumerate}
\end{comment}

\bibliographystyle{dcu}
\bibliography{includes/bibliography}

\end{document}